---
title: "Lab3B"
author: "Julia Sousa"
date: "9/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}

## load the tidyverse packages, include dplyr

library("tidyverse")

```


# What are dplyr and tidyr?


```{r}

surveys <- read_csv("portal_data_joined.csv")

```

```{r}

## inspect the data
str(surveys)

```


# Preview the data

```{r}
view(surveys)

```



# Selecting columns and flitering rows

```{r}

select(surveys, plot_id, species_id, weight)

select(surveys, -record_id, -species_id)

filter(surveys, year == 1995)

```

# Pipes


```{r}

surveys2 <- filter(surveys, weight <5)
surveys_sml <- select (surveys2, species_id, sex, weight)

```


```{r}

surveys_sml <- select(filter(surveys, weight <5,), species_id, sex, weight)

```


```{r}

surveys %>% 
  filter(weight < 5) %>% 
  select(species_id, sex, weight)

```


```{r}

surveys_sml <- surveys %>% 
  filter(weight < 5) %>% 
  select(species_id, sex, weight)
surveys_sml

```



# Challenge 1

```{r}

surveys %>% 
  filter(year > 1995) %>% 
  select(year, sex, weight)

```


# Mutate

```{r}

surveys %>% 
  mutate(weight_kg = weight / 1000)

```


```{r}

surveys %>% 
  mutate(weight_kg = weight / 1000,
         weight_lb = weight_kg * 2.2)

```


```{r}

surveys %>% 
  mutate(weight_kg = weight / 1000) %>% 
  head()

```



```{r}

surveys %>% 
  filter(!is.na(weight)) %>% 
  mutate(weight_kg = weight / 1000) %>% 
  head()

```



# Challenge 2

```{r}

surveys_hindfoot_cm <- surveys %>%
  filter(!is.na(hindfoot_length)) %>% 
  mutate(hindfoot_cm = hindfoot_length / 10) %>% 
  filter(hindfoot_cm <3) %>% 
  select(species_id, hindfoot_cm)

surveys_hindfoot_cm

```


# Split-apply-combine data analysis and the summarize() function

```{r}

surveys %>% 
  group_by(sex) %>% 
  summarize(mean_weight = mean(weight, na.rm = TRUE))

```


#>`summarise()' ungrouping output (override with `.groups` argument)



```{r}

surveys %>% 
  group_by(sex, species_id) %>% 
  summarize(mean_weight = mean(weight, na.rm = TRUE)) %>% 
  tail()

```



#>`summarise()` regrouping output by `sex` (override with `.groups` argument)


```{r}

surveys %>% 
  filter(!is.na(weight)) %>% 
  group_by(sex, species_id) %>% 
  summarize(mean_weight = mean(weight))

```


#>`summarise()` regrouping output by `sex` (override with `.groups` argument)

```{r}

surveys %>% 
  filter(!is.na(weight)) %>% 
  group_by(sex, species_id) %>% 
  summarize(mean_weight = mean(weight)) %>% 
  print(n = 15)

```


#>`summarise()` regrouping output by `sex` (override with `.groups` argument)

```{r}

surveys %>% 
  filter(!is.na(weight)) %>% 
  group_by(sex, species_id) %>% 
  summarize(mean_weight = mean(weight),
            min_weight = min(weight))

```

#>`summarise()` regrouping output by `sex` (override with `.groups` argument)

```{r}

surveys %>% 
  filter(!is.na(weight)) %>% 
  group_by(sex, species_id) %>% 
  summarise(mean_weight = mean(weight),
            min_weight = min(weight)) %>% 
  arrange(min_weight)

```

#>`summarise()` regrouping output by `sex` (override with `.groups` argument)

```{r}

surveys %>% 
  filter(!is.na(weight)) %>% 
  group_by(sex, species_id) %>% 
  summarize(mean_weight = mean(weight),
            min_weight = min(weight)) %>% 
  arrange(desc(mean_weight))

```


#>`summarise()` regrouping output by `sex` (override with `.groups` argument)


```{r}

surveys %>% 
  count(sex)

```


```{r}

surveys %>% 
  group_by(sex) %>% 
  summarise(count = n())

```


#>`summarise()` regrouping output by `sex` (override with `.groups` argument)


```{r}

surveys %>% 
  count(sex, sort = TRUE)

```

```{r}

surveys %>% 
  count(sex, species)
```

```{r}
surveys %>% 
  count(sex, species) %>% 
  arrange(species, desc(n))

```


# Challenge 3


```{r}

surveys %>% 
  count(plot_type)

```

```{r}

surveys %>% 
  filter(!is.na(hindfoot_length)) %>% 
  group_by(species_id) %>% 
  summarize(mean_hindfoot_length = mean(hindfoot_length),
            min_hindfoot_length = min(hindfoot_length),
            max_hindfoot_length = max(hindfoot_length),
            n =n()
  )

```
#>`summarise()` regrouping output by `sex` (override with `.groups` argument)


```{r}

surveys %>% 
  filter(!is.na(weight)) %>% 
  group_by(year) %>% 
  filter(weight == max(weight)) %>% 
  select(year, genus, species, weight) %>% 
  arrange(year)

```


# Reshaping with gather and spread


```{r}

surveys_gw <- surveys %>% 
  filter(!is.na(weight)) %>% 
  group_by(plot_id, genus) %>% 
  summarize(mean_weight = mean(weight))

#> `summarise()` regrouping by 'plot_id' (override with `.groups` arguement)

str(surveys_gw)

```


```{r}

surveys_spread <- surveys_gw %>% 
  spread(key = genus, value = mean_weight)

str(surveys_spread)

```

```{r}

surveys_gw %>% 
  spread(genus, mean_weight, fill =0) %>% 
  head()

```


```{r}

surveys_gather <- surveys_spread %>% 
  gather(key = "genus", value = "mean_weight", -plot_id)

str(surveys_gather)

```


```{r}

surveys_spread %>% 
  gather(key = "genus", value = "mean_weight", Baiomys:Spermophilus) %>% 
  head()

```


# Challenge 4

```{r}

surveys_spread_genera <- surveys %>% 
  group_by(plot_id, year) %>% 
  summarize(n_genera = n_distinct(genus)) %>% 
  spread(year, n_genera)

#> `summarise()` regrouping output by 'plot_id' (override with `.groups` argument)

head(surveys_spread_genera)

```


```{r}

surveys_long <- surveys %>% 
  gather("measurement", "value", hindfoot_length, weight)
surveys_long

```


```{r}

surveys_long %>% 
  group_by(year, measurement, plot_type) %>% 
  summarize(mean_value = mean(value, na.rm=TRUE)) %>% 
  spread(measurement, mean_value)

#> `summarise()` regrouping output by 'year', 'measurement' (override with `.groups` argument)

```



# Exporting data

```{r}

surveys_complete <- surveys %>% 
  filter(!is.na(weight),            # remove missing weight
         !is.na(hindfoot_length),      # remove missing hindfoot_length
         !is.na(sex))               # remove missing sex

```


```{r}

## Extract the most common species_id
species_counts <- surveys_complete %>% 
  count(species_id) %>% 
  filter(n >= 50)

## Only keep the most common species
surveys_complete <- surveys_complete %>% 
  filter(species_id %in% species_counts$species_id)

```


```{r}

dim(surveys_complete)

```

```{r}

write_csv(surveys_complete, path = "data/surveys_complete.csv")

```




