---
title: "Lab 9"
author: "Julia Sousa"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 14.3: Matching patterns with regular expresions

```{r}

library(tidyverse)
library(stringr)

```


# 14.3.1 Basic matches

```{r}

x <- c("apple", "banana", "pear")
str_view(x, "an")

```


```{r}

str_view(x, ".a.")

```


```{r}

dot <- "\\."

writeLines(dot)

```

```{r}

x <- c("abc", "a.c", "bef")
str_view(x, "a\\.c")

```

```{r}

x<- "a\\b"

writeLines(x)

```

```{r}
str_view(x, "\\\\")

```

# Exercise 1 (14.3.1.1)

If "\" is used, it will escape the next character in the regular expression.
If "\\" is used, it will mean "\" in the regular expression.
If "\\\" is used, the first two backslashes will equal "\" in the regular expression, and the third backslash will escape the next character in the regular expression.



# 14.3.2. Anchors

```{r}

x <- c("apple", "banana", "pear")
str_view(x, "^a")

```

```{r}

str_view(x, "a$")

```

```{r}

x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")

```

```{r}

str_view(x, "^apple$")

```

# Exercise 1 (14.3.2.1)

You would match the literal string "$^$" by:

```{r}

str_view("$^$", "ab$^$sfas")

```


## 14.3.3 Character classes and alternatives

```{r}

x <- c("abc", "a.c", "a*c", "a c")
str_view(x, "a[.]c")

```

```{r}

str_view(x, ".[*]c")

```

```{r}

str_view(x, "a[ ]")

```

```{r}

x <- c("grey", "gray")
str_view(x, "gr(e|a)y")

```

# Exercise 1 (14.3.3.1)

Words that start with a vowel:
```{r}

str_subset(stringr::words, "^[aeiou]")

```

Words that only contain consonants:
```{r}

str_subset(stringr::words, "[aeiou]", negate=TRUE)

```

Words that end with "ed", but not in "eed":
```{r}

str_subset(stringr::words, "[^ed]ed$")

```

Words ending in "ing" or "ise":
```{r}

str_subset(stringr::words, "i(ng|se)$")

```


## 14.3.4 Repitition

```{r}

x <- "188 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")

```

```{r}

str_view(x, "CC+")

```

```{r}

str_view(x, 'C[LX]+')

```

```{r}

str_view(x, "C{2}")

```

```{r}

str_view(x, "C{2,}")

```

```{r}

str_view(x, "C{2,}")

```

```{r}

str_view(x, "C{2,3}?")

```

```{r}

str_view(x, "C[LX]+?")

```

# Exercise 1 (14.3.4.1)

? is equivalent to {0,1}
+ is equivalent to {1,}
* is equivalent to {0,}



## 14.3.5 Grouping and backreferences



















