---
title: "Lab 9"
author: "Julia Sousa"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 14.3: Matching patterns with regular expresions

```{r}

library(tidyverse)
library(stringr)

```


# 14.3.1 Basic matches

```{r}

x <- c("apple", "banana", "pear")
str_view(x, "an")

```


```{r}

str_view(x, ".a.")

```


```{r}

dot <- "\\."

writeLines(dot)

```

```{r}

x <- c("abc", "a.c", "bef")
str_view(x, "a\\.c")

```

```{r}

x<- "a\\b"

writeLines(x)

```

```{r}
str_view(x, "\\\\")

```

# Exercise 1 (14.3.1.1)

If "\" is used, it will escape the next character in the regular expression.
If "\\" is used, it will mean "\" in the regular expression.
If "\\\" is used, the first two backslashes will equal "\" in the regular expression, and the third backslash will escape the next character in the regular expression.



# 14.3.2. Anchors

```{r}

x <- c("apple", "banana", "pear")
str_view(x, "^a")

```

```{r}

str_view(x, "a$")

```

```{r}

x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")

```

```{r}

str_view(x, "^apple$")

```

# Exercise 1 (14.3.2.1)

You would match the literal string "$^$" by:

```{r}

str_view("$^$", "ab$^$sfas")

```


## 14.3.3 Character classes and alternatives

```{r}

x <- c("abc", "a.c", "a*c", "a c")
str_view(x, "a[.]c")

```

```{r}

str_view(x, ".[*]c")

```

```{r}

str_view(x, "a[ ]")

```

```{r}

x <- c("grey", "gray")
str_view(x, "gr(e|a)y")

```

# Exercise 1 (14.3.3.1)

Words that start with a vowel:
```{r}

str_subset(stringr::words, "^[aeiou]")

```

Words that only contain consonants:
```{r}

str_subset(stringr::words, "[aeiou]", negate=TRUE)

```

Words that end with "ed", but not in "eed":
```{r}

str_subset(stringr::words, "[^ed]ed$")

```

Words ending in "ing" or "ise":
```{r}

str_subset(stringr::words, "i(ng|se)$")

```


## 14.3.4 Repitition

```{r}

x <- "188 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")

```

```{r}

str_view(x, "CC+")

```

```{r}

str_view(x, 'C[LX]+')

```

```{r}

str_view(x, "C{2}")

```

```{r}

str_view(x, "C{2,}")

```

```{r}

str_view(x, "C{2,}")

```

```{r}

str_view(x, "C{2,3}?")

```

```{r}

str_view(x, "C[LX]+?")

```

# Exercise 1 (14.3.4.1)

? is equivalent to {0,1}
+ is equivalent to {1,}
* is equivalent to {0,}



## 14.3.5 Grouping and backreferences

```{r}

str_view(fruit, "(..)\\1", match = TRUE)


```

# Exercise 1 (14.3.5.1)

(.)\1\1 will match the same character that appears three times in a row

"(.)(.)\\2\\1" will match a pair of characters followed by the same pair of characters, but in a reversed order

(..)\1 will match any two characters repeated

"(.).\\1.\\1" will match a character followed by another character, the first character, another character, and the first character again

"(.)(.)(.).*\\3\\2\\1" will match three characters followed by zero or more characters of any kind, followed by the same three characters, but in a reversed order


## 14.4.1 Detech matches

```{r}

x <- c("apple", "banana", "pear")
str_detect(x, "e")

```
```{r}

sum(str_detect(words, "^t"))

```

```{r}

mean(str_detect(words, "[aeoiu]$"))

```

```{r}

no_vowels_1 <- !str_detect(words, "[aeiou]")

```


```{r}

no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)

```


```{r}

words[str_detect(words, "x$")]

```

```{r}

str_subset(words, "x$")

```

```{r}

df <- tibble(
  word = words,
  i = seq_along(word)
)
df %>% 
  filter(str_detect(word, "x$"))

```

```{r}

x <- c("apple", "banana", "pear")
str_count(x, "a")

```

```{r}

mean(str_count(words, "[aeiou]"))

```

```{r}

df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )

```

```{r}

str_count("abababa", "aba")

```

```{r}

str_view_all("abababa", "aba")

```

# Exercise 1 (14.4.1.1.)

Words that start or end with x:

```{r}

words[str_detect(words, "^x|x$")]

```

Words that start with a vowel and end with a consonant:

```{r}

vowel_starting <- str_detect(words, "^[aeiou]")
consonant_ending <- str_detect(words, "[^aeiou]$")
words[vowel_starting & consonant_ending]

```

Words that contain at least one of each different vowel:

```{r}

words[str_detect(words, "a") &
        str_detect(words, "e") &
        str_detect(words, "i") &
        str_detect(words, "o") &
        str_detect(words, "u")
      ]

```

## 14.4.2 Extract matches

```{r}

length(sentences)

```

```{r}

head(sentences)

```

```{r}

colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match

```

```{r}

has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)

```

```{r}

has_colour

```

```{r}

more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)

```

```{r}

str_extract(more, colour_match)

```

```{r}

str_extract_all(more, colour_match)

```

```{r}

str_extract_all(more, colour_match, simplify = TRUE)

```

```{r}

x <- c("a", "a b", "a b c")
str_extract_all(x, "[a-z]", simplify= TRUE)

```

# Exercise 1 (14.4.2.1)

```{r}

colour_match2 <- str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")
colour_match2

```
```{r}

more2 <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more2, colour_match2, match = TRUE)

```


## 14.4.3 Grouped matches

```{r}

noun <- "(a|the) ([^ ]+)"

has_noun <- sentences %>% 
  str_subset(noun) %>% 
  head(10)
has_noun %>% 
  str_extract(noun)

```

```{r}

has_noun %>% 
  str_match(noun)

```

```{r}

tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)",
    remove = FALSE
  )

```


# Exercise 1 (14.4.3.1)

Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.

```{r}

number_word <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten) +(\\w+)"
sentences[str_detect(sentences, number_word)] %>% 
  str_extract(number_word)

```

## 14.4.4 Replacing matches

```{r}

x <- c("apple", "banana", "pear")
str_replace(x, "[aeiou]", "-")

```

```{r}

str_replace_all(x, "[aeiou]", "-")

```

```{r}

x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))

```

```{r}

sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)

```

## Exercise 1 (14.4.4.1)

```{r}

str_replace_all("small/medium/large", "/", "\\\\")

```

14.4.5 Splitting

```{r}

sentences %>% 
  head(5) %>% 
  str_split(" ")

```

```{r}

"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]

```

```{r}

sentences %>% 
  head(5) %>% 
  str_split(" ", simplify = TRUE)

```


```{r}

fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(":", n = 2, simplify = TRUE)

```

```{r}

x <- "This is a sentence. This is another sentence."
str_view_all(x, boundary("word"))

```

```{r}

str_split(x, " ")[[1]]

```

```{r}

str_split(x, boundary("word"))[[1]]

```

## Exercise 1 (14.4.5.1)















